---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Sandy Yu
thanks: "Code and data are available at: https://github.com/Jingying-yu/help-line-homeless-death-analysis.git"
date: today
date-format: long
abstract: "This paper analyzes the influence of Central Intake Line staff efforts on the reduction of homeless death count in Toronto. Data on coded number of calls from Central Intake Call Wrap-Up Codes dataset and data on monthly homeless death count from Deaths of People Experiencing Homelessness dataset are used in generalized linear regression models to estimate the influence. We find that the staffs’ net efforts create a statistically insignificant influence on the reduction of homeless death, with specific types of efforts appearing counterproductive. This implies that while a positive influence does exist, from a statistical point of view, the impact of the Central Intake Line staff is not significantly different from non-existent."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(rstanarm)
library(arrow)
library(knitr)
library(ggplot2)
```

# Introduction

The City of Toronto has been consistently providing funding and staff in an effort to reduce the active homeless population within its governing district. Amongst numerous city efforts, the 24/7 helpline — Toronto Central Intake Line — was created to offer referrals to emergency shelters, sleeping accommodations, and provide general information about homelessness and prevention. [@Central_Intake_Call]. Starting in 2020, detailed data about the Central Intake Line became available on the OpenDataToronto portal [@OpenDataToronto].

The estimand of the paper focuses on examining the influence of helpline efforts on the reduction of homeless deaths in Toronto. 
Using Central Intake Call Wrap-up Codes data [@Central_Intake_Call] from OpenDataToronto [@OpenDataToronto], three different aspects of the helpline efforts were isolated: total calls coded, referral to shelter, and homelessness prevention information provided. These three key aspects were measured against the monthly cumulative death count from Deaths of People Experiencing Homelessness Data [@Homeless_Death_Count] documented by the Toronto Public Health [@TPH]. The analyses yielded an insignificant negative correlation between the net efforts of the helpline staff. The variable referral to shelter yields a positive correlation with homeless death count, whereas the effort to provide homelessness information yields a negative correlation. This implies that the net influence of helpline efforts is statistically negligible, while individual efforts can be productive or counterproductive. This examination in the net efforts and individual efforts can provide the helpline with a suggestion of which types of services or calls would be more economically efficient to allocate their labor resources on.

Analyses and findings in this paper are structured into several sections: @sec-data -- Data, @sec-model – Model, @sec-results -- Results, and @sec-discussion -- Discussion. The Data section examines all datasets and variables kept for analysis, followed by an explanation of their data cleaning processes. The Model section defines linear models used for further analysis, explain its components, and presents model justifications. The Result section focuses on visualizing and presenting the model results through data presented in Data section. The Discussion section further evaluate the interpretations behind the model results presented in the previous section, and touches on any weaknesses and next steps.

# Data {#sec-data}

All data used in this paper are obtained through OpenDataToronto Portal [@OpenDataToronto]. Two different datasets: Central Intake Call Wrap-Up Codes Data [@Central_Intake_Call] and Deaths of People Experiencing Homelessness[@Homeless_Death_Count], are retrieved to analyze the effect of Toronto’s Central Call Line (denoted as *Helpline* for the remainder of this paper) efforts to the death counts of homeless individuals in Toronto. Data is cleaned and analyzed using the open source statistical programming language R [@citeR] and supporting packages tidyverse [@tidyverse], janitor [@janitor], dplyr [@dplyr], rstanarm [@rstanarm], arrow [@arrow], ggplot2 [@ggplot2], and knitr [@knitr]. Detailed description of each dataset can be found in the subsections below.

## Central Intake Call Wrap-Up Codes

On the OpenDataToronto portal, there are several datasets that reflect the City’s effort to shelter the local homeless population. The Central Intake Call Wrap-Up Codes Dataset [@Central_Intake_Call], stored in the Central Intake Calls Catalogue [@Central_Intake_Call], is one of the freshest and most detailed. Data is stored and published by the Shelter, Support & Housing Administration since November of 2020 and refreshes on a monthly basis. The latest refresh occurred on January 15th, 2024.

The data set provides a daily summary of the number of calls received, the number of calls classified into distinct wrap-up codes by the nature of its issue, and a count of calls under each wrap-up code. One of the example wrap-up codes in the original data set was: *Code 1A - Referral to a Sleeping/Resting Space*. The original data set includes 13 distinct wrap-up codes; only two codes, *Code 1A — Referral to a Sleeping/Resting Space* and *Code 2C — Information - Homelessness & Prevention Services*, were chosen for our analysis. Code 1A and Code 2C are best suited as measurements for positive impact the Central Intake Line could provide because they provide a count of the number of callers provided with directions or advice. Other non-suitable wrap-up codes include: *Code 1D - Declined Shelter/Resting Space*, *Code 4B - Disconnected - No Outcome*, etc.

The final dataset only includes monthly cumulative data before July of 2023. The rationalization behind the action will be explained in @sec-combined-data.

```{r}
#| label: tbl-monthly-coded
#| tbl-cap: Sample of monthly summation for Helpline Coded Calls data
#| echo: false

cleaned_data <- read_parquet(here::here("data/analysis_data/cleaned_data.parquet"))

sample_helpline <- cleaned_data |> select(Month_date, Total_Coded, Referred, Informed) 

sample_helpline <- head(sample_helpline)

knitr::kable(sample_helpline, col.names = c("Month", "Calls Coded", "Referral to Shelter", "Homelessness Prevention Info"), align = c("l", "c", "c", "c"))
```

## Deaths of People Experiencing Homelessness

The Deaths of People Experiencing Homelessness Dataset [@Homeless_Death_Count] contains monthly cumulative records of homeless deaths. The dataset is published by Toronto Public Health [@TPH]. The earliest data record started in January of 2017, and the latest record ends in June of 2023.

The original dataset contains three columns: *Year of death*, *Month of death*, and *Count*. After careful inspection of raw data, any rows that containing an “unknown” value are excluded. Although this action created a source of error in future analyses, this is still a necessary step because we are plotting death counts against a timeline. 

Final dataset only includes data from November 2020 to June 2023 (@tbl-death-count). The rationalization behind the action will be explained in @sec-data-diff-datasets.

```{r}
#| label: tbl-death-count
#| tbl-cap: Sample of Cleaned Deaths of People Experiencing Homelessness Dataset
#| echo: false

sample_death_counts <- cleaned_data |> select(Month_date, Death_Count) 

sample_death_counts <- head(sample_death_counts)

knitr::kable(sample_death_counts, col.names = c("Month", "Death Count"), align = c("c", "c"))
```

## Final Combined Data {#sec-combined-data}

During the data cleaning process, code was written to ensure that both the Central Intake Call Wrap-Up Codes Dataset [@Central_Intake_Call] and the Deaths of People Experiencing Homelessness Dataset [@Homeless_Death_Count] are filtered to include only data between 1st November, 2020 to 30st June, 2023. 

The decision is reached by taking the common time period between the two datasets. In the latest refresh, the Central Intake Call Wrap-Up Codes Dataset [@Central_Intake_Call] begins on 3rd November, 2020 and end on 31st December, 2023; whereas the Deaths of People Experiencing Homelessness Dataset [@Homeless_Death_Count] begins on January of 2017 and ends on June 2023.

The ultimate purpose of taking the common time period is to ensure that we can have the same amount of observations for analysis. Thus a combined dataset that kept the variables of interest in both datasets is created (@tbl-combined).

```{r}
#| label: tbl-combined
#| tbl-cap: Sample of monthly homeless death count and Helpline Coding efforts
#| echo: false

sample_cleaned_data <- head(cleaned_data)

knitr::kable(sample_cleaned_data, col.names = c("Month", "Total Coded", "Referred to Shelter", "Homeless Info", "Death Count"), align = c("l", "c", "c", "c", "c"))

```

# Model {#sec-model}

While we can still make claims on the correlations between helpline efforts and homeless death count without modelling, regression analysis is still necessary to provide rigorous backbones for a claim's statistical significance.

Here we briefly describe the Bayesian analysis model used to investigate the correlation between homeless death counts in Toronto and the efforts of the Helpline staff. Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Magnitude of effect created by the helpline efforts can be estimated through linear models. Gaussian regression model is best suited for our analytical purposes. Other generalized linear models such as Logistic and Multilevel are not considered due to the nature of the variables in interest. Logistic regression model requires data that yields either an **yes** or **no** result, whereas multilevel model require more complexity between variables. Due to the nature of the Poisson function, the Poisson model was considered and discarded because model results in extremely small values with no evidently different interpretations.

### Model 1: Total Coded Model

$y_i|\mu_i$, $\sigma \sim \mbox{Normal}(\mu_i, \sigma)$

$mu_i = \alpha + \phi_i$

$\alpha \sim \mbox{Normal}(0, 2.5)$

$\phi \sim \mbox{Normal}(0, 2.5)$

$\sigma \sim \mbox{Exponential}(1)$

Where:

-   $y_i$ is the number of homeless death count in Toronto recorded by @TPH per month

-   $\phi_i$ is the total number of calls received and coded by the helpline per month

### Model 2: Referred Informed Model

$y_i|\mu_i$, $\sigma \sim \mbox{Normal}(\mu_i, \sigma)$

$\mu_i = \alpha + \beta_i + \gamma_i$

$\alpha \sim \mbox{Normal}(0, 2.5)$

$\beta \sim \mbox{Normal}(0, 2.5)$

$\gamma \sim \mbox{Normal}(0, 2.5)$

$\sigma \sim \mbox{Exponential}(1)$

Where:

-   $y_i$ is the number of homeless death count in Toronto recorded by @TPH per month

-   $\beta_i$ is the number of homeless individuals referred to shelters through the helpline, per month

-   $\gamma_i$ is the number of individuals provided with information on homelessness prevention through the helpline, per month

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justifications

We expect a net negative relationship between the net efforts made by the helpline staff (Recorded by the Total_Coded) and homeless death count, which implies that net effect of helpline calls **reduces** homeless death count in Toronto.

Additionally, we expect a negative relationship between homeless death and the two individual efforts: *number of homeless individuals referred to shelter*, and *number of individuals provided with homelessness prevention information*. Since both factors are coded efforts made by the helpline, both factors should also help to reduce homeless death count in Toronto.

# Results {#sec-results}

Our results are summarized in @tbl-total_coded_modelresults and @tbl-referred_informed_modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

total_coded_model <-
  readRDS(file = here::here("models/total_coded_model.rds"))

referred_informed_model <-
  readRDS(file = 
            here::here("models/referred_informed_model.rds"))
```

```{r}
#| label: fig-total_coded
#| fig-cap: Visualization to show correlation between number of Calls Coded and Homeless Death Counts
#| echo: false
#| message: false

cleaned_data |> ggplot(aes(x = Total_Coded, y = Death_Count)) + geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm", se = TRUE, color = "blue", linetype = "dashed", formula = "y ~ x") +
  labs(x = "Monthly Number of Calls Coded (per month)", y = "Monthly Number of Homeless Deaths")
```


```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory model of homeless death count in relation to different helpline efforts"
#| tbl-subcap: "statistical significances are indicated by p < 0.1 *, p < 0.05 **, p < 0.01 ***, p < 0.001"
#| warning: false

modelsummary::modelsummary(
  list(
    "Total Coded Model" = total_coded_model, "Referred and Informed Model" = referred_informed_model
  ),
  statistic = "mad",
  stars = TRUE,
  fmt = 4
)
```




# Discussion {#sec-discussion}

## First discussion point {#sec-first-point}


## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-total_coded_model_ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-total_coded_model_ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-total_coded_model_ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(total_coded_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(total_coded_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

In @fig-referred_informed_model_ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-referred_informed_model_ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-referred_informed_model_ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(referred_informed_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(referred_informed_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-total_coded_model_stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-total_coded_model_stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-total_coded_model_stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm for total_coded_model"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(total_coded_model, "trace")

plot(total_coded_model, "rhat")
```

@fig-referred_informed_model_stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-referred_informed_model_stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-referred_informed_model_stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm for referred_informed_model"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(referred_informed_model, "trace")

plot(referred_informed_model, "rhat")
```

\newpage

# References
